---
title: "IS624 Week 7"
author: "Aaron Palumbo"
date: "Monday, July 20, 2015"
output: pdf_document
toc: true
---

## Question 1: Plotting/Comparing Survival Curves Part I

Load in the ovarian data-set:

```{r}
library(survival)
data(ovarian)
head(ovarian)
```

This data is from a randomised trial comparing two treatments for ovarian cancer. Look in http://cran.r-project.org/web/packages/survival/survival.pdf, the variables are as follows:  

> **futime:** survival or censoring time  
> **fustat:** censoring status  
> **age:** in years  
> **resid.ds:** residual disease present (1-no, 2-yes)  
> **rx:** treatment group  
> **ecog.ps:** ECOG performance status (1 is better)  

Compare the two treatment groups (**treatment1** versus **treatment2**). Plot the respective survival curves, indicating censored subjects. You can distinguish between the two groups using different colors or different line formats or both. Label both x- and y-axes. Add a suitable title. Also, add a legend indicating which line corresponds to which line format.

Finally compare the two survival curves (log-rank) and add a p-value to the bottom-right of the plot.

HINT: You add tow or more legends to the plot; to add a legend without a border use box.col="white".

You should achieve something like this:

```{r fig.height=5}
fit <- survfit(Surv(futime, fustat)~rx, data=ovarian)
survdiff(Surv(futime,fustat)~rx, data=ovarian)

plot(fit, main="OVARIAN CANCER - OVERALL SURVIVAL",
     xlab="Time (Days)", ylab="Overall Survival Proportion", col.lab='blue',
     col=c("red", "blue"), lty=2:3)
legend("topright", lty=2:3, col=c("red", "blue"),
       legend=c("Treatment Group1",
                "Treatment Group2"))
legend(x=800, y=.2, legend="p-value: 0.303", box.col="white")

```


## Question 2: Plotting/Comparing Survival Curves Part

Load in the Leukaemia-free survival/transplant data-set:

```{r}
library(KMsurv)
data(alloauto)
head(alloauto)
```

This data considers tow transplant types in relation to leukaemia-ftee survival. Look in http://cran.r-project.org/web/packages/KMsurv/KMsurv.pdf, the variables are as follows:

> **Time:** Time to death or relapse, months  
> **Type:** Type of transplant (1-allogeneic, 2-autologous)  
> **Delta:** Leukemia-free survival indicator(0=alive withoutrelapse, 1-deador relapse)  

Compare the two transplant types (**allogeneic** vs **autologous**). Plot the respective relapse-free survival curves, indicating censored subjects. You can distinguish between the two groups using different colors or different line formats or both. Label both x- and y-axes. Add a suitable title. Also, add a legend indicating which line corrresponds to which line format.

Finally compare the two survival curves (log-rank) and add a p-value to the bottom-right of the plot.

You should achieve somthing like this:

```{r fig.height=5}
fit <- survfit(Surv(time, delta)~type, data=alloauto)
survdiff(Surv(time, delta)~type, data=alloauto)

plot(fit, main="Leukemia - Free Survival",
     xlab="Time (Months)", ylab="Relapse-Free Survival Proportion", col.lab='blue',
     col=c("red", "blue"), lty=2:3)
legend(40, 0.9, lty=2:3, col=c("red", "blue"),
       legend=c("Allogeneic", "Autologous"), cex=0.8)
text(30, .2, labels="p-value: 0.537", cex=0.8)
```


## Question 3: Developing a Cox Model

Look at the hmoiv data-set again.

Recode the age variables (see below) and repeat the univariate and multivariate analyses.

```{r}
library(car)
hmohiv <- read.table("http://www.ats.ucla.edu/stat/R/examples/asa/hmohiv.csv", 
                     sep=",", header = TRUE) 
```

### Univariate

```{r}
agecat <- recode(hmohiv$age, "20:29='D'; 30-34='B'; 35:39='C'; 40:54='A'", as.factor=TRUE)

coxph(formula=Surv(time, censor)~agecat, method="efron", data=hmohiv)
```








