---
title: "IS624 Week 7"
author: "Aaron Palumbo"
date: "Monday, July 20, 2015"
output: pdf_document
---

Question 1: Plotting/Comparing Survival Curves Part I

Load in the ovarian data-set:

```{r}
library(survival)
data(ovarian)
ovarian
```

This data is from a randomised trial comparing two treatments for ovarian cancer. Look in http://cran.r-project.org/web/packages/survival/survival.pdf, the variables are as follows:  

> **futime:** survival or censoring time  
> **fustat:** censoring status  
> **age:** in years  
> **resid.ds:** residual disease present (1-no, 2-yes)  
> **rx:** treatment group  
> **ecog.ps:** ECOG performance status (1 is better)  

Compare the two treatment groups (**treatment1** versus **treatment2**). Plot the respective survival curves, indicating censored subjects. You can distinguish between the two groups using different colors or different line formats or both. Label both x- and y-axes. Add a suitable title. Also, add a legend indicating which line corresponds to which line format.

Finally compar the tow survival curves (log-rank) and add a p-value to the bottom-right of the plot.

HINT: You add tow or more legends to the plot; to add a legend without a border use box.col="white".

You should achieve something like this:

```{r}
gr1Filter <- ovarian$rx == 1
survG1 <- Surv(ovarian$futime[gr1Filter], ovarian$fustat[gr1Filter])
fitG1 <- survfit(survG1 ~ 1, conf.type="none")

survG2 <- Surv(ovarian$futime[-gr1Filter], ovarian$fustat[-gr1Filter])
fitG2 <- survfit(survG2 ~ 1, conf.type="none")
plot(fitG1)
lines(fitG2)
```

